---
title: "Lab 1: Group GA"
output: pdf_document
date: "2022-09-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW_1
Exercise 1: Sum up all even numbers from 10 to 1002 (look at the help page of sum() and seq())
```{r}

n<-seq(10, 1002, 2)
n
sum(n)

```

Exercise 2: Create a vector of strings where the elements are "Hello","world!", "How", "are", and "you?". Combine all of these words together into one string, separated by underscores (look at help page of paste())

```{r}
v<-c("Hello","world!", "How", "are", "you?")
output<-paste(v, collapse = "_")
output
```
Exercise 3:  What happend when you sum "1" with 2?
```{r}
#whathappned<-sum("1", 2) # two type of variable(charcter, double)
#whathappned  
#...Error in sum("1", 2) : invalid 'type' (character) of argument
```


Exercise 4: You can comment out code using hashtags # at the start of a line. That is, the code will no longer run.Comments are great to make code more readable. Comment out the code in the previous chunk and rerun.
```{r}
#whathappned<-sum("1", 2) # two type of variable(charcter, double)
#whathappned  
```
Exercise 5: If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Find the sum of all the multiples of 3 or 5 below 1000.

```{r}
answer=0
for (i in 1:999) {
    if (i%%3 == 0 | i%%5 == 0) 
        answer <- answer + i}
answer
```
Exercise 6: (Extra Credit)What is the sum of all integers that are either (divisible by 4 and less than 700) or (divisible by 3 and between
500 and 1000)?

```{r}
answer=0
for (i in 1:1000) {
    if  ((i%%4 == 0 & i < 700) |(i%%3==0 & i > 500 & i< 1000))
        answer <- answer + i}
answer
```

