---
title: "HW_2_Team_GA"
output: pdf_document
date: "2022-09-16"
---

```{r setup, include= FALSE }
knitr::opts_chunk$set(echo = TRUE)
```
> ### Instructions

* After completing the questions, upload both the .RMD and PDF files to Bb.

* Learning Outcomes:

* Create functions

* Create and source scripts

* Employ piping

* Employ logicals

* Employ conditionals.
     
>###  1. Implement a fizzbuzz() function
    
    1. Create a new function which takes a single number as input. Use logicals and conditionals as well as the operator %% as required.
    
• Requirements

– If the input number is divisible by three, return "fizz".

– If it’s divisible by five, return "buzz".

– If it’s divisible by three and five, return "fizzbuzz".

– Otherwise, return the input number.

```{r }
fizzbuzz <- function(single_number){
  if(single_number%%5==0 & single_number%%3==0)
    return("fizzbuzz")
  else if(single_number%%5==0)
    return("buzz")
  else if (single_number%%3==0)
    return("fizz")
  else
    return(single_number)}


```
  
     a. Design your function. Write out in words in a text chunk the steps your function will need to accomplishin order to convert the input number into the required text or numeric output.
     
1. use %% == 0, so we can set the divisor.

2. assign the function to take "single_number"

3. make sure all arguments are in {} in the right position.

4. set the number divisible 3 and 5 to retunr “fizzbuzz” 

5. set the number divisible 3 to return  “fizz”

6. set the number divisible 5 to return “buzz”

7. otherwise, I set to retunr the "single_number" itself

8. use else if function 


    b. Write R code in a code chunk to implement your steps using a variable x. Test it with x having values of of 3, 5, 15, and 16.
    
```{r }
fizzbuzz <- function(x){
  if(x%%5==0 & x%%3==0)
    return("fizzbuzz")
  else if(x%%5==0)
    return("buzz")
  else if (x%%3==0)
    return("fizz")
  else
    return(x)}

```
```{r}
fizzbuzz(3)
```

```{r}
fizzbuzz(5)
```

```{r}
fizzbuzz(15)
``` 
    
```{r}
fizzbuzz(16)
```   
    c. Once the code is working, then copy it to a new code chunk and turn it into function with input argument of x.
    
```{r}
fizzbuzz <- function(x){
  if(x%%5==0 & x%%3==0)
    return("fizzbuzz")
  else if(x%%5==0)
    return("buzz")
  else if (x%%3==0)
    return("fizz")
  else
    return(x)}
```

    d. Show your output for the following inputs: 3, 5, 15, 2.
    
```{r }
fizzbuzz(3)
```
```{r }
fizzbuzz(5)
```
```{r }
fizzbuzz(15)
```
```{r }
fizzbuzz(2)
```

    e. Update your function to include error checking
    
• Ensure the input is both numeric and a single value - not a vector.

• Test it on cat, and c(1,5).

• Remember, in the code chunk where you run the function, set your code chunk parameter for
error to be TRUE,

– e.g. {r, error=TRUE}, so it will knit with the error.

```{r, error=TRUE }
fizzbuzz <- function(x) {
  stopifnot(length(x) == 1) # checking if x is single input, if not error
  stopifnot(is.numeric(x)) # Checking if x is numeric, if not error  
 if(x%%5==0 & x%%3==0)
    return("fizzbuzz")
  else if(x%%5==0)
    return("buzz")
  else if (x%%3==0)
    return("fizz")
  else
    return (as.character(x))}

fizzbuzz('cat')

```

```{r, error=TRUE }
fizzbuzz <- function(x) {
  stopifnot(length(x) == 1) # checking if x is single input, if not error
  stopifnot(is.numeric(x)) # Checking if x is numeric, if not error  
 if(x%%5==0 & x%%3==0)
    return("fizzbuzz")
  else if(x%%5==0)
    return("buzz")
  else if (x%%3==0)
    return("fizz")
  else
    return (as.character(x))}

fizzbuzz(c(1,5))

```

  f. Complete your function by inserting and completing Roxygen comments in the code chunk, above the function, to document the function.Include the following elements: title, description, usage or syntax, arguments ( the params), and return value.

```{r, error=TRUE }
# Implement a fizzbuzz() function which takes a single number as input
# x: is a single number as input
# If the number is divisible by three and five, return "fizzbuzz"
# If the number is divisible by three, return "fizz"
# If the number is divisible by five, return "buzz"


fizzbuzz <- function(x) {
  stopifnot(length(x) == 1) # checking if x is single input, if not error
  stopifnot(is.numeric(x)) # Checking if x is numeric, if not error  
 if(x%%5==0 & x%%3==0)
    return("fizzbuzz")
  else if(x%%5==0)
    return("buzz")
  else if (x%%3==0)
    return("fizz")
  else
    return (as.character(x))}

```    
    
    g. Create a script out of your fizzbuzz() function
    
• Copy and paste the code from your working function into a new .R file and save in the R directory
with the file name fizzbuzz_s.R

• Rename the function to fizzbuzz_s

• Use the following code in a code chunk to show your code

– cat(readr::read_file("./R/fizzbuzz_s.R"))

– Adjust the relative path as necessary

• Write code in a new code chunk in your original homework file to source the fizzbuzz_s() function

• Run the function in your homework .Rmd file to show the results with the values 35, 18, 45, and -1

```{r}
cat(readr::read_file("./fizzbuzz_s.R"))
```
```{r}
source("~/Desktop/Desktop - FENTAW’s MacBook Air/American_U/R_programming/fizzbuzz_s.R")
fizzbuzz_s(35)
```


```{r}
fizzbuzz_s(18)

```

```{r}
fizzbuzz_s(45)

```

```{r}
fizzbuzz_s(-1)

```

>### 2. Create a new cut() function
    
    1. Write a function that uses the function cut() to simplify this set of nested if-else statements?
    
  • Consider using -Inf and Inf.
  
  • Note, this will also output the levels of the factors.
      
      a. Show the output for inputs: 31, 30, 10, -10.

```{r}
# #temp=c(-Inf,0,10,20,30,Inf)
# if (temp <= 0)
#   return("freezing")
# else (temp <= 10)
#   return("cold")
# else (temp <= 20)
#   return("cool")
# } else if (temp <= 30) {
# "warm"
# } else {
# "hot"}
```

```{r}
temp_type <- function(temp){
  seq(-10,50,by = 5)
cut(temp, c(-Inf,0,10,20,30,Inf),right = TRUE,
labels = c("freezing", "cold", "cool", "warm", "hot"))}
```
```{r}
temp_type(31)
```
```{r}
temp_type(30)
```    
    
```{r}
temp_type(10)
```    
   
```{r}
temp_type(-10)
```   
   
    b. Look at help for cut(). Change the call to cut() to handle < instead of <= in the comparisons.

```{r}
?cut()

# cut(x, ...)
# 
# ## Default S3 method:
# cut(x, breaks, labels = NULL,
#     include.lowest = FALSE, right = TRUE, dig.lab = 3,
#     ordered_result = FALSE, ...)
# cut divides the range of x into intervals and codes the values in x according to which interval they fall. The leftmost interval corresponds to level one, the next leftmost to level two and so on.
temp_type <- function(temp){
  seq(-10,50,by = 5)
cut(temp, c(-Inf,0,10,20,30,Inf),right = FALSE,
labels = c("freezing", "cold", "cool", "warm", "hot"))}

temp_type(30)
```

    c. What is the other chief advantage of the cut() method for this problem? (Hint: what happens if you have many values in temp?)

```{r}
# By using cut, I can works on vectors,and 
# To change comparisons I only needed to change the argument to "right"
```

>### 3. Using the Forward Pipe
    
    1. Using the forward pipe %>%, 
    
  • Sample from the vector 1:10 1000 times with replacement,
  
  • Calculate the resulting sampled vector’s mean, then
  
  • Exponent that mean.
  

```{r}
set.seed(123)
library(dplyr)
```
```{r}
library(magrittr)
library(knitr)
sample(c(1:10),1000,TRUE) %>% mean() %>%exp()
```
>### 4. Calculate a proportion

• Select a random sample of 100 normally distributed values with mean 10 and variance of 3.

• Calculate the proportion greater than 12.

```{r}
library(magrittr)
sd1=sqrt(3)
x<- rnorm(100,mean=10,sd=sd1)
proportion_greater12= mean(x>12) # use the mean function to roll this up to a proportion
proportion_greater12


### Note: every time I run the chunk, the proportion output changed because of rnorm. 
```

```{r}
### This is the other way I've tried, 
sd1=sqrt(3) # sd is a sqrt of variance.
x<- rnorm(100,mean=10,sd=sd1)
pnorm(12,mean=10,sd=sd1, lower.tail = TRUE)
```



>###  5. Logical Comparisons and Subsetting

• Create the values:

– x <- c(TRUE, FALSE, TRUE, TRUE)

– y <- c(FALSE, FALSE, TRUE, FALSE)

– z <- NA

• What are the results of the following:

– x & y

```{r}
x <- c(TRUE, FALSE, TRUE, TRUE)

y <- c(FALSE, FALSE, TRUE, FALSE)

z <- NA

x&y
```

– x & z
 
```{r}
x&z
```

– !(x | y)

```{r}
!(x | y)
```


– x | y

```{r}
x | y
```

– y | z

```{r}
y | z
```

– x[y]

```{r}
x[y]
```

– y[x]

```{r}
y[x]
```

– x[x|y]
   
```{r}
x[x|y]
```

