---
title: "HW_3_Team_GA"
output: pdf_document
date: "2022-09-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, header= TRUE)
```

### Use the following libraries in order to write code and execute output. 
 
 library(ggplot2) library(tidyverse) 

Show and use R coding to answer the following questions. ( Use Tidyverse methods to 
generate  graphs and plots)

1) Explore the Midwest data frame.    

```{r }
library(tidyverse) 
library(dplyr)
library(ggplot2)
data("midwest")
```
a) How many rows, columns, and variables are in the Midwest data frame? 

```{r }
glimpse(midwest)
?midwest
```
Rows: 437
Columns: 28


b) Name three categorical variables in the data frame. 

```{r }
split(names(midwest),sapply(midwest, function(x) paste(class(x), collapse=" ")))
```
Categorical variable:
1. "county": County name   
2. "state": State to which county belongs to.
3. "category": Miscellaneous 
4. "inmetro": County considered in a metro area 


c) Give a description for the variable percollege. 


```{r }
library(psych)
describe(midwest$percollege)

```
percollege: Percent college educated.


Use the Midwest data frame for problems 2,3,4,5,6,7, and 9


### 2) Write R code to produce a histogram for the variable poptotal.


```{r }
library(ggplot2) 
bw <- 2 * IQR(midwest$poptotal) / length(midwest$poptotal)^(1/3)  # Freedman–Diaconis rule
ggplot(midwest, aes(x=poptotal))+
  geom_histogram(binwidth= bw, bins = sqrt(437)) 
# Based on the rule, 
#it isn't much clear; if we set binwidth as such.
```


```{r }
library(ggplot2) 
library(tidyverse)
ggplot(midwest, aes(x=poptotal))+
  geom_histogram(bins = sqrt(437)) #Better look without the binwidth set.

      
```

### 3) Write r code to produce a boxplot for the variable area, and then use your box plot to find Q1, Q2, and Q3 

```{r}
boxplot(midwest$area)
```
```{r}
library(mosaic)
favstats(midwest$area)
```

### 4) Write r code to produce side by side boxplots for the quantitative variable area with respect to the categorical variable state. 



```{r}
library(ggplot2)
midwest$state <-factor(midwest$state) # converts state to a categorical variable
my.area_state <<-ggplot(data=midwest, aes(y=area, x=state, fill=state ) ) # Creates boxplots
my.area_state <- my.area_state + geom_boxplot()
my.area_state
```

### 5) Write r code to produce a scatter plot for the variables popdensity and popwhite. Let popdensity be the independent variable x and popwhite be dependent variable y.

```{r}
plot_scatter <- ggplot(midwest) + geom_point(aes(y = popdensity , x = popwhite))
plot_scatter
```

### 6) Write r code that will produce smooth lines plots and scatter plots on the same axis system for popwhite and popdensity with respect to the categorical variable state. 

```{r}
library(ggplot2)
plot_scatter_smooth <- ggplot(midwest,aes(y = popdensity , x = popwhite, color = state)) + 
  geom_point() +
  geom_smooth(se = FALSE) #geom_smooth left to default which set ` using method = 'loess' and formula 'y ~ x'

plot_scatter_smooth

#Note: we are not asked to do lm here; just smoothing lines 
```

### 7) Again,  using the variables popdensity and popwhite, write r code that will produce the same basic scatter plot, but also make the following changes: 


a) Your scatter plot should have purple data points. 

```{r}
plot_scatter <- ggplot(midwest, aes(y = popdensity , x = popwhite)) + 
  geom_point(color= "purple")
plot_scatter
```

b) The label of the dependent variable should be changed to Population  Density and the 
label of the independent variable should be changed to White Population

```{r}
plot_scatter_level <- ggplot(midwest, aes(y = popdensity , x = popwhite)) + 
  geom_point(color= "purple") +
  xlab("White Population") +
  ylab("Population Desnsity")
plot_scatter_level
```

c) Add a linear regression line to your graph. 

```{r}
plot_scatter_level_smoothing <- ggplot(midwest, aes(y = popdensity , x = popwhite)) + 
  geom_point(color= "purple") +
  xlab("White Population") +
  ylab("Population Desnsity") +
  geom_smooth(se=FALSE, method = lm)
plot_scatter_level_smoothing
```
d) Add the following title to your graph; White vs Density Scatter Plot 

```{r }
plot_scatter_level_smoothing <- ggplot(midwest, aes(y = popdensity , x = popwhite)) + 
  geom_point(color= "purple") +
  xlab("White Population") +
  ylab("Population Desnsity") +
  ggtitle("White vs Density Scatter Plot")+
  geom_smooth(se=FALSE, method = lm)
plot_scatter_level_smoothing
```

### 8) Write R code that will generate the following graph (use the mpg data frame) 

```{r }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
geom_smooth(se = TRUE)
```

### 9   Write R code that will produce the following facet plot using the midwest data frame

```{r}
ggplot(data = midwest, mapping = aes(x = inmetro, y = popdensity)) +
  geom_point() +
  facet_wrap( ~ state)
```

### 10   Write R code that will produce the following facet plot using the mpg data frame
ggplot(data = mpg, aes(x=cty)) +
  geom_histogram(position="identity", fill="green", alpha=1, bins = 30) +
  facet_wrap( drv ~cyl )
```{r}
ggplot(data = mpg, aes(x=cty)) +
  geom_histogram(fill="green", alpha=0.5) +
  facet_grid(drv ~cyl)

```
Note: the Alpha  =0.5  and the bin set to the default is an estimation because its hard to know the density of the color and size of the bin from your graph.
