---
title: "HW_6_R"
author: "Group_GA"
date: "2022-10-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of tidyr is to help you create tidy data. Tidy data is data where:
Every column is variable.
Every row is an observation.
Every cell is a single value.


# Exercise 1
Tidy the data frame ex0724 from the Sleuth3 package. You can read about this data frame by typing
help(ex0724) after loading Sleuth3.

```{r}

library(tidyverse)
library(dplyr)
library(tidyr)
```

```{r}
Sleuth3::ex0724
```
```{r}
library(Sleuth3)
data("ex0724") 
library(magrittr)
library(readr)

```



```{r}
ex0724 %>%
  tidyr::gather(Denmark:USA, key = "country", value = "proportion", na.rm = TRUE) -> clean_ex0724
```

```{r}
clean_ex0724
```

# Exercise 2

Load in and tidy the tb data frame: https://dcgerard.github.io/stat_412_612/data/tb.csv
The column names specify both the sex (m = male, f = female) and age range (04 = 0 to 4, 514 = 5 to 14,
014 = 0 to 14, 1524 = 15 to 24, 2534 = 25 to 34, 3544 = 35 to 44, 4554 = 45 to 54, 4464 = 55 to 64, 65 =>=65, u = unknown).

The values in the cells are counts.Save the tidied data in the output folder.


```{r}
tb<-read_csv("https://dcgerard.github.io/stat_412_612/data/tb.csv") 
tb
```

```{r}
tb %>%
  tidyr::gather(-iso2, -year, key = "sex_age", value = "counts", na.rm = TRUE) %>%
  tidyr::separate(col = sex_age, into = c("sex", "age"), sep = 1) ->
  tb2
tb2
```


# Exercise 3

Load in and tidy the wine data frame: https://dcgerard.github.io/stat_412_612/data/wine.csv
Save the tidied data in the output folder.

```{r}
wine <- read_csv2("https://dcgerard.github.io/stat_412_612/data/wine.csv")
wine
```
```{r}
wine %>%
  tidyr::gather(-measure, key = "country", value = "value") %>%
  tidyr::spread(key = measure, value = value) ->
  wine_clean
wine_clean
```

