---
title: "HW_Last"
author: "GA"
date: "2022-11-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Learning Objectives
•Practice stringr, dplyr, and ggplot2.
•WARNING: Do not use str_view() or str_view_all() on these data. It will stall your computer.
The data aren’t that large, but str_view() and str_view_all() are inefficient with medium data.
– More stringr options can be found in RDS.
# Exercise 1: From RDS:

1. Replace all forward slashes in a string with backslashes. Test it out on the following string:

x <- "hello\\////how//are////you////"

```{r}
library(tidyverse)
x <- "hello\\////how//are////you////"
y <- str_replace_all(x, "/", "\\\\")
writeLines(x)
writeLines(y)
```


2. Construct regular expressions to match words that:

a. Start and end with the same character. A word of length 1 should be matched. Test it out on "A",
"AB", and "ABA".




```{r}
str_detect("A", "^((.).*\\2|.)$")
str_detect("AB", "^((.).*\\2|.)$")
str_detect("ABA", "^((.).*\\2|.)$")
```

b. Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.) Test it out on "AAA",
"AAAA", and "AABAA".

```{r}
str_detect("AAA", "(..).*\\1")
str_detect("AAAA", "(..).*\\1")
str_detect("AABAA", "(..).*\\1")
```

c. Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.) Test it out on
"AAA", "AAB", and "AABA".


```{r}
str_detect("AAA", "(.).*\\1.*\\1")
str_detect("AAB", "(.).*\\1.*\\1") 
str_detect("AABA", "(.).*\\1.*\\1")
```


FACTORS
# Exercise 2: Sex, Lies, and Religion
The data frame in https://dcgerard.github.io/stat_412_612/data/sexlierel.txt, taken from @clay-
ton1971religiosity, contain the following variables:


•gender: The gender of the individual. 1 = Female, 2 = Male.
•scale: Different scales of sexual permissiveness. 1 = Ritualistic, 2 = Experiential, 3 = Ideological, 4
= Composite.
•perm: The level of sexual permissiveness. 1 = Low, 2 = High.
•lie: The propensity to lie. 1 = Lower, 2 = Higher.
relig: How religious a person is. 1 = Low, 2 = High
•count: The number of individuals satisfying the conditions of the other variables.

1. Read the data into R.

```{r}
library(tidyverse)
library('httr')
sexlierel <- read_table(file = "https://dcgerard.github.io/stat_412_612/data/sexlierel.txt", 
                            col_types = cols(
                              gender = col_factor(),
                              scale  = col_factor(),
                              perm   = col_factor(),
                              lie    = col_factor(),
                              relig  = col_factor(),
                              count  = col_double()
                              ))
```

2. Change the level names to something more informative.

```{r}
sexlierel %>%
  mutate(gender = fct_recode(gender, 
                             "Female" = "1", 
                             "Male"   = "2"),
         
         scale  = fct_recode(scale,
                             "ritualistic"  = "1", 
                             "experiential" = "2",
                             "idiological"  = "3",
                             "composite" = "4"), 
         perm   = fct_recode(perm,
                             "low"  = "1", 
                             "high" = "2"), 
         lie    = fct_recode(lie,
                             "lower"  = "1",
                             "higher" = "2"),
         relig  = fct_recode(relig, 
                             "low"  = "1", 
                             "high" = "2")) ->sexlierel

head(sexlierel)
```

3. Flip the order of the levels in perm.

```{r}
 sexlierel %>%
      mutate(perm = fct_rev(perm)) ->
      sexlierel
head(sexlierel)
```

4. For males, for what scales does there appear to be an association between religiosity and permissiveness?
Use one plot to explore.

```{r}
 sexlierel %>%
      filter(gender == "Male") %>%
      ggplot(aes(x = relig, y = count, fill = perm)) +
      geom_col(pos = "fill") +
      facet_grid(. ~ scale) +
      ylab("Proportion") +
      xlab("Religiosity") +
      scale_fill_discrete(name = "Permissiveness") +
      theme(strip.background = element_rect(fill = "white"))

```
 for all


